---
title: 'Data Challenge: Exploratory Data Analysis'
author: "Aldo Iturrios"
date: "3/27/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(stringr)
```

```{r, include = FALSE}
# Data
clean_data <- read.csv("../Data/clean_data/clean_data.csv")
train_data <- read.csv("../Data/clean_data/train_data.csv")
test_data <- read.csv("../Data/clean_data/test_data.csv")
```

```{r, include = FALSE}
time_var_subset <- grep(paste("[0-9]{2}.[0-9]{2}_.", sep=""), colnames(clean_data))
time_data <- clean_data[,time_var_subset]
var_names <- unique(names(time_data[ ,6:ncol(time_data)]))
timestamps <- c(unique(str_extract(var_names, "[0-9]{2}.[0-9]{2}")))
measures <- unique(str_extract(var_names, "(?<=_)[a-zA-Z]+[0-9]*"))
```


## NA's in the Data

### Question: What percentage of patients have at least one record for a particular variable (ie. 60% of patients have at least 1 blood pressure reading)

```{r, include = FALSE}
nas_vec <- c()
for (i in 1:length(measures)) {
  if (measures[i] == "SAPS") { 
    nas_vec[i] <- sum(is.na(clean_data[,grep(measures[i], colnames(clean_data))]) > 0) / nrow(clean_data)
  }
  else {
    nas_vec_var <- rowSums(is.na(clean_data[,grep(measures[i], colnames(clean_data))]))
    nas_vec[i] <- sum((ncol(clean_data[,grep(measures[i], colnames(clean_data))]) - nas_vec_var) > 0) / nrow(clean_data)
  }
}
```


```{r, echo=FALSE}
data.frame(measures, patients_atleast_one = round(nas_vec * 100, 2))
```

### Question: Distribution of patient data at each time point (ie. 98% of patients have data at 00:30, 80% of patients have data at 06:30, etc…)

```{r, include = FALSE}
times_vec <- c()
for (i in 1:length(timestamps)) {
  times_vec_var <- rowSums(is.na(clean_data[,grep(timestamps[i], colnames(clean_data))]))
    times_vec[i] <- sum((ncol(clean_data[,grep(timestamps[i], colnames(clean_data))]) - times_vec_var) > 0) / nrow(clean_data)
}
```

```{r}
data.frame(timestamps, patients_atleast_one = round(times_vec * 100, 2))
```

# EDA on Training Data (Looking at variables in relation to Outcome)

### Outcome:

```{r}
# Training Data
table(train_data$outcome)
```

### Recovery by Admission Type

```{r}
counts <- table(train_data$outcome, train_data$AdmissionType)
barplot(counts, main="Recovery by Admission Type", xlab="Admission Type", ylab = "Number of Patients", col=c("darkblue","red"), legend = rownames(counts), beside=TRUE)
addmargins(counts)
```

### Recovery by Gender

```{r}
gender_counts <- table(train_data$outcome, train_data$Gender)
barplot(gender_counts, main="Recovery by Gender", xlab="Gender", ylab = "Number of Patients", col=c("darkblue","red"), legend = rownames(gender_counts), beside=TRUE)
addmargins(gender_counts)
```

```{r}
summary(train_data$Age)
```

```{r}
boxplot(train_data$Age ~ train_data$outcome)
```

# Look at a few measure summary variables

```{r, fig.show="hold", out.height = "50%", out.width="50%", echo=FALSE}
hist(train_data$ALP_mean)
hist(train_data$ALP_median)
```

```{r, fig.show="hold", out.height = "50%", out.width="50%", echo=FALSE}
hist(train_data$Glucose_mean)
hist(train_data$Glucose_median)
```

```{r, fig.show="hold", out.height = "50%", out.width="50%", echo=FALSE}
hist(train_data$Cholesterol_mean)
hist(train_data$Cholesterol_median)
```

```{r, fig.show="hold", out.height = "50%", out.width="50%", echo=FALSE}
hist(train_data$Urine_mean)
hist(train_data$Urine_median)
```


